---
title: "Factor Analysis"
author: "pg606@scarletmail.rutgers.edu"
date: "2024-03-19"
output: html_document
---


```{r }
library(psych)
```


```{r }
data <- read.csv("/Users/Prateekg/Downloads/data1.csv", row.names=1)
attach(data)
data[1]
```


```{r }
fit.pc <- principal(data[-1], nfactors=4, rotate="varimax")
fit.pc
round(fit.pc$values, 3)
fit.pc$loadings
```

```{r }
for (i in c(1,3,2,4)) { print(fit.pc$loadings[[1,i]])}
```



```{r }
fit.pc$communality
```



```{r }
fit.pc$scores
fa.parallel(data[-1]) 
```

# Question1: Decide how many Factors are ideal for your dataset?
# In this case, the output suggests that the parallel analysis suggests that the number of factors is 1, and the number of components is 1. However, based on the initial factor analysis performed, it seems like four factors (RC1, RC2, RC3, RC4) have been extracted.

# Question 2: Explain the output for your factor model?

#•	Standardized loadings (pattern matrix) based upon the correlation matrix: This table shows the relationships between the original variables and the extracted factors. Each cell represents the correlation between a variable and a factor.

#•	SS loadings: This indicates the sum of squared loadings for each factor, providing information about how much variance in the original variables is explained by each factor.

#•	Proportion Var: This represents the proportion of variance explained by each factor.

#•	Cumulative Var: It shows the cumulative proportion of variance explained by the factors.

#•	Proportion Explained: This shows the proportion of total variance explained by each factor.

#•	Cumulative Proportion: It indicates the cumulative proportion of total variance explained by the factors.

#•	Mean item complexity: This indicates the average number of parameters required to describe an item.

#•	Test of the hypothesis that 4 components are sufficient: This section provides statistical tests to determine the adequacy of the chosen number of factors.

#•	Root mean square of the residuals (RMSR) and empirical chi-square: These are measures of how well the model fits the data.


# Question 3: Show the columns that go into each factor?

#RC1: Salary experience ratio and years of experience.

#RC2: Salary level.

#RC3: Location diversity.

#RC4: Salary in USD.


#Question 4: Perform some visualizations using the factors?

#Below are the Visualizations-




```{r }
fa.plot(fit.pc) 
fa.diagram(fit.pc) 
vss(data[-1]) 

```



```{r }

```


```{r }

```


```{r }

```


```{r }

```



```{r }

```



```{r }

```


```{r }

```



```{r }

```



```{r }

```

