---
title: "LDA_Salary"
author: "pg606@scarletmail.rutgers.edu"
date: "2024-04-26"
output: html_document
---

```{r }
library(MASS)
library(ggplot2)
library(memisc)
library(ROCR)
library(dplyr)
library(klaR)

wdbc <- read.csv("/Users/Prateekg/Documents/MVA/Salary_multiple_regression.csv")
dim(wdbc)
str(wdbc)
features <- c("Category", "Currency", "EmpResidence","ExpLevel", "EmpType", "WorkSetting", "location", "size", "level", "year", "salary", "USD", "SalaryLevel")
names(wdbc) <- c("Title", "Category", "Currency", "EmpResidence","ExpLevel", "EmpType", "WorkSetting", "location", "size", "level", "year", "salary", "USD", "SalaryLevel")

```

#Model development

```{r }
wdbc.data <- as.matrix(wdbc[,c(11:13)])
row.names(wdbc.data) <- wdbc$Title
wdbc_raw <- cbind(wdbc.data, as.numeric(as.factor(wdbc$SalaryLevel))-1)
colnames(wdbc_raw)[4] <- "EmpSalaryLevel"
smp_size_raw <- floor(0.75 * nrow(wdbc_raw))
train_ind_raw <- sample(nrow(wdbc_raw), size = smp_size_raw)
train_raw.df <- as.data.frame(wdbc_raw[train_ind_raw, ])
test_raw.df <- as.data.frame(wdbc_raw[-train_ind_raw, ])
wdbc_raw.lda <- lda(formula = train_raw.df$EmpSalaryLevel ~ ., data = train_raw.df)
wdbc_raw.lda

```

Split the dataset into training and testing sets.
Examine the output of the trained LDA model to understand the prior probabilities of groups, group means, and coefficients of linear discriminants.
Interpret the coefficients to understand the impact of each predictor on the salary level prediction.


#Model Acceptance


```{r }
summary(wdbc_raw.lda)
print(wdbc_raw.lda)
plot(wdbc_raw.lda)

```

Group means indicate the average values of predictor variables for each salary level class.
For class 0 (low salary), the average year is 2016, salary is $60,000, and USD is $60,000.
For class 1 (high salary), the average year is 2018, salary is $265,800, and USD is $265,800.
These differences suggest notable variations in salary-related variables between the two salary levels.
#Accuracy

```{r }
wdbc_raw.lda.predict_train <- predict(wdbc_raw.lda, newdata = train_raw.df)
y<-wdbc_raw.lda.predict_train$class
wdbc_raw.lda.predict_train$x
table(y,train_raw.df$EmpSalaryLevel)

```




```{r }

wdbc_raw.lda.predict <- predict(wdbc_raw.lda, newdata = test_raw.df)
y<-wdbc_raw.lda.predict$class
wdbc_raw.lda.predict$x
table(y,test_raw.df$EmpSalaryLevel)
```

Both the training and test set accuracies are 100%, indicating that the model perfectly classified all instances in both datasets.


#Residual Analysis

```{r }
# Get the posteriors as a dataframe.
wdbc_raw.lda.predict.posteriors <- as.data.frame(wdbc_raw.lda.predict$posterior)
```

Residual analysis to refine the model, such as adjusting predictor variables, exploring different modeling techniques, or collecting additional data.


#Prediction
```{r }
pred <- prediction(wdbc_raw.lda.predict.posteriors[1:3,2], test_raw.df$EmpSalaryLevel)


roc.perf = performance(pred, measure = "tpr", x.measure = "fpr")
auc.train <- performance(pred, measure = "auc")
auc.train <- auc.train@y.values
plot(roc.perf)
abline(a=0, b= 1)
text(x = .25, y = .65 ,paste("AUC = ", round(auc.train[[1]],3), sep = ""))

```


```{r }
plot(wdbc_raw.lda, col = as.integer(train_raw.df$EmpSalaryLevel))

plot(wdbc_raw.lda, dimen = 1, type = "b")

m <- manova(cbind(wdbc$year,wdbc$salary,wdbc$USD)~wdbc$SalaryLevel,data=wdbc)
summary(m,test="Wilks")

summary(m,test="Pillai")

summary.aov(m)
```

The ROC curve and AUC value provide insights into the discriminative performance of the LDA model. A higher AUC suggests better separation between the two salary levels.
The plots of the LDA model visualize how well the model separates the data points belonging to different salary levels.
The MANOVA results indicate whether there are significant differences in predictor variables across different salary levels. Significant p-values suggest that at least one predictor variable differs significantly between salary levels.


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```


```{r }

```

```{r }

```

```{r }

```


```{r }

```



```{r }

```
