---
title: "Principle Component Analysis"
author: "pg606@scarletmail.rutgers.edu"
date: "2024-03-04"
output: html_document
---

```{r }
library(readr)
dataset <- read.csv("/Users/Prateekg/Downloads/Dataset_with_Salary_Level.csv")
dataset
str(dataset)
attach(dataset)

```

```{r }
cor(dataset[,10:13][-1])

```

```{r }
dataset_pca <- prcomp(dataset[,10:13][,-1],scale=TRUE)
dataset_pca
summary(dataset_pca)
```

```{r }
(eigen_dataset <- dataset_pca$sdev^2)
names(eigen_dataset) <- paste("PC",1:3,sep="")
eigen_dataset
sumlambdas <- sum(eigen_dataset)
sumlambdas
propvar <- eigen_dataset/sumlambdas
propvar
cumvar_dataset <- cumsum(propvar)
cumvar_dataset
matlambdas <- rbind(eigen_dataset,propvar,cumvar_dataset)
rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
round(matlambdas,4)
summary(dataset_pca)
dataset_pca$rotation
print(dataset_pca)
```

QUESTION : Decide how many Principal Components (PCs) you want to keep and why?






ANSWER : To decide how many principal components (PCs) to keep, we typically consider the proportion of variance explained by each PC. In this case, we have three principal components, and their proportions of variance are as follows:

PC1: 0.6808
PC2: 0.3163
PC3: 0.00289
The cumulative proportion of variance is also provided:

PC1: 0.6808
PC2: 0.9971
PC3: 1.0000
From this information, we can see that PC1 explains 68.08% of the variance, PC2 explains 31.63% of the variance, and PC3 explains only 0.289% of the variance.

Given that PC1 already explains a significant portion of the variance (68.08%) and PC2 contributes substantially as well (31.63%), you might decide to keep the first two principal components. PC3 contributes very little to the overall variance.

Therefore, a reasonable choice would be to keep the first two principal components, PC1 and PC2, as they collectively explain 99.71% of the variance. Keeping these two components would retain most of the variability in the original dataset while reducing dimensionality.







QUESTION 2 : Explain the variate representation each PCs


ANSWER 2 : PC1:

PC1 has the highest standard deviation and explains the largest proportion of variance (68.08%).
The variables with the most significant influence on PC1 are salary and salary_in_usd, both negatively correlated. This suggests that individuals with higher salaries tend to have lower values for salary_in_usd and vice versa.
Work_year also contributes to PC1 but to a lesser extent. The negative loading indicates that as work_year increases, the value of PC1 decreases.
Overall, PC1 captures the variation related to the salary and its representation in USD, possibly reflecting different salary structures or exchange rates.
PC2:

PC2 has a lower standard deviation compared to PC1 but still explains a substantial proportion of variance (31.63%).
Work_year has the most dominant influence on PC2 with a strong negative loading. This suggests that as work_year increases, the value of PC2 decreases significantly.
Salary and salary_in_usd also contribute to PC2, but with weaker positive loadings compared to their influence on PC1. This indicates that while higher salaries are associated with higher values of PC2, the effect is not as pronounced as with work_year.
PC2 seems to capture the variation related to work experience (work_year) primarily, with some additional influence from salary-related variables.
PC3:

PC3 has the lowest standard deviation and explains a negligible proportion of variance (0.29%).
All variables contribute minimally to PC3. Work_year and salary have very small loadings, while salary_in_usd has a slightly higher loading but still very close to zero.
PC3 does not provide meaningful information in explaining the variance within the dataset. It captures residual variance not accounted for by PC1 and PC2.
It is common for the last principal component to capture noise or unexplained variance in the data.
In summary, PC1 primarily represents the variation in salary and its representation in USD, PC2 captures variation related to work experience, and PC3 captures residual variance. These interpretations provide insights into the underlying structure of the dataset and can guide further analysis or modeling efforts.
```{r }
dataset_pca$x
# Identifying the scores by their salary status
datasettyp_pca <- cbind(data.frame(Salary_Level),dataset_pca$x)
datasettyp_pca

tabmeansPC <- aggregate(datasettyp_pca[,2:4],by=list(Salary_Level=dataset$Salary_Level),mean)
tabmeansPC
tabmeansPC <- tabmeansPC[rev(order(tabmeansPC$Salary_Level)),]
tabmeansPC
tabfmeans <- t(tabmeansPC[,-1])
tabfmeans
colnames(tabfmeans) <- t(as.vector(tabmeansPC[1]$Salary_Level))
tabfmeans
# Standard deviations of scores for all the PC's classified by Survival status
tabsdsPC <- aggregate(datasettyp_pca[,2:4],by=list(Salary_Level=dataset$Salary_Level),sd)
tabfsds <- t(tabsdsPC[,-1])
colnames(tabfsds) <- t(as.vector(tabsdsPC[1]$Salary_Level))
tabfsds
```



```{r }
t.test(PC1~dataset$Salary_Level,data=datasettyp_pca)
t.test(PC2~dataset$Salary_Level,data=datasettyp_pca)
t.test(PC3~dataset$Salary_Level,data=datasettyp_pca)

var.test(PC1~dataset$Salary_Level,data=datasettyp_pca)
var.test(PC2~dataset$Salary_Level,data=datasettyp_pca)
var.test(PC3~dataset$Salary_Level,data=datasettyp_pca)
```

```{r }
library(car)
(LTPC1 <- leveneTest(PC1~dataset$Salary_Level,data=datasettyp_pca))
(p_PC1_1sided <- LTPC1[[3]][1]/2)
(LTPC2 <- leveneTest(PC2~dataset$Salary_Level,data=datasettyp_pca))
(p_PC2_1sided=LTPC2[[3]][1]/2)
(LTPC3 <- leveneTest(PC3~dataset$Salary_Level,data=datasettyp_pca))
(p_PC3_1sided <- LTPC3[[3]][1]/2)
```

```{r }
plot(datasettyp_pca$PC1, datasettyp_pca$PC2,pch=ifelse(datasettyp_pca$Salary_Level == "Yes",1,16),xlab="PC1", ylab="PC2", main="Dataset against values for PC1 & PC2")
abline(h=0)
abline(v=0)
legend("bottomleft", legend=c("Dataset","No Dataset"), pch=c(1,16))
plot(eigen_dataset, xlab = "Component number", ylab = "Component variance", type = "l", main = "Scree diagram")
plot(log(eigen_dataset), xlab = "Component number",ylab = "log(Component variance)", type="l",main = "Log(eigenvalue) diagram")
print(summary(dataset_pca))
diag(cov(dataset_pca$x))
xlim <- range(dataset_pca$x[,1])
dataset_pca$x[,1]
dataset_pca$x
plot(dataset_pca$x,xlim=xlim,ylim=xlim)
dataset_pca$rotation[,1]
dataset_pca$rotation
plot(dataset[,-1])
dataset_pca$x
plot(dataset_pca)
#get the original value of the data based on PCA
center <- dataset_pca$center
scale <- dataset_pca$scale
new_dataset <- as.matrix(dataset[,10:13][,-1])
new_dataset
drop(scale(new_dataset,center=center, scale=scale)%*%dataset_pca$rotation[,1])
predict(dataset_pca)[,1]
#The aboved two gives us the same thing. predict is a good function to know.
dataset$Salary_Level <- as.factor(dataset$Salary_Level)
out <- sapply(1:3, function(i){plot(dataset$Salary_Level,dataset_pca$x[,i],xlab=paste("PC",i,sep=""),ylab="dataset")})
pairs(dataset_pca$x[,1:3], ylim = c(-6,4),xlim = c(-6,4),panel=function(x,y,...){text(x,y,dataset$Salary_Level)})

```

VISUALIZATION :

```{r }
library(factoextra)
library(FactoMineR)
library(ggfortify)
library(psych)
library(corrplot)
library(devtools)




# Correlation
pairs.panels(dataset[,-1],
             gap = 0,
             bg = c("red", "blue")[dataset$Salary_Level],
             pch=21)


pairs.panels(dataset_pca$x,
             gap=0,
             bg = c("red", "blue")[dataset$Salary_Level],
             pch=21)

```


```{r }
fviz_eig(dataset_pca, addlabels = TRUE)



fviz_pca_var(dataset_pca,col.var = "cos2",
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"),
             repel = TRUE)
fviz_pca_ind(dataset_pca, col.ind = "cos2", 
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"), 
             repel = TRUE)
```


```{r }
biplot(dataset_pca)
autoplot(dataset_pca,
         data = dataset[,-1],
         loadings = TRUE,
         labels = dataset$Salary_Level)

# Different PCA Method. 
res.pca <- PCA(dataset[,10:13][,-1], graph = FALSE)
print(res.pca)

# Visualize and Interpret PCA using these functions 

#get_eigenvalue(res.pca): Extract the eigenvalues/variances of principal components
#fviz_eig(res.pca): Visualize the eigenvalues
#get_pca_ind(res.pca), get_pca_var(res.pca): Extract the results for individuals and variables, respectively.
#fviz_pca_ind(res.pca), fviz_pca_var(res.pca): Visualize the results individuals and variables, respectively.
#fviz_pca_biplot(res.pca): Make a biplot of individuals and variables.

eig.val <- get_eigenvalue(res.pca)
eig.val

fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))

var <- get_pca_var(res.pca)
```

```{r }
#The plot Below is also known as variable correlation plots. It shows the relationships between all variables. It can be interpreted as follow:

#Positively correlated variables are grouped together.
#Negatively correlated variables are positioned on opposite sides of the plot origin (opposed quadrants).
#The distance between variables and the origin measures the quality of the variables on the factor map. 
#Variables that are away from the origin are well represented on the factor map.

# Correlation circle
fviz_pca_var(res.pca, col.var = "black")
```

```{r }
# Quality of representation


corrplot(var$cos2, is.corr=FALSE)
# Total cos2 of variables on Dim.1 and Dim.2
#A high cos2 indicates a good representation of the variable on the principal component. 
#In this case the variable is positioned close to the circumference of the correlation circle.
#A low cos2 indicates that the variable is not perfectly represented by the PCs. 
#In this case the variable is close to the center of the circle.

fviz_cos2(res.pca, choice = "var", axes = 1:2)
fviz_pca_var(res.pca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE # Avoid text overlapping
)
```

```{r }
# Change the transparency by cos2 values
fviz_pca_var(res.pca, alpha.var = "cos2")
corrplot(var$contrib, is.corr=FALSE)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
fviz_pca_var(res.pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
)
```

```{r }
fviz_pca_var(res.pca, alpha.var = "contrib")

fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = dataset$Salary_Level, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
)
```

```{r }

```


```{r }
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = dataset$Salary_Level, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
)
```


```{r }

fviz_pca_ind(res.pca, geom.ind = "point", col.ind = dataset$Salary_Level, 
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, ellipse.type = "confidence",
             legend.title = "Groups"
)

```

```{r }
fviz_pca_ind(res.pca,
             label = "none", # hide individual labels
             habillage = dataset$Salary_Level, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             palette = "jco"
)

```

```{r }
fviz_pca_var(res.pca, geom.var = c("point", "text"))
# Show individuals text labels only
fviz_pca_ind(res.pca, geom.ind =  "text")
# Change the size of arrows an labels
fviz_pca_var(res.pca, arrowsize = 1, labelsize = 5, 
             repel = TRUE)


```

```{r }
# Change points size, shape and fill color
# Change labelsize
fviz_pca_ind(res.pca, 
             pointsize = 3, pointshape = 21, fill = "lightblue",
             labelsize = 5, repel = TRUE)

fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (but not "text")
             group.ind = dataset$Salary_Level, # color by groups
             legend.title = "Groups",
             mean.point = FALSE)
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (but not "text")
             group.ind = dataset$Salary_Level, # color by groups
             legend.title = "Groups",
             mean.point = TRUE)
fviz_pca_var(res.pca, axes.linetype = "blank")
```

```{r }
ind.p <- fviz_pca_ind(res.pca, geom = "point", col.ind = dataset$Salary_Level)
ggpubr::ggpar(ind.p,
              title = "Principal Component Analysis",
              subtitle = "dataset data set",
              caption = "Source: factoextra",
              xlab = "PC1", ylab = "PC2",
              legend.title = "Salary Level", legend.position = "top",
              ggtheme = theme_gray(), palette = "jco"
)
```




```{r }
fviz_pca_biplot(res.pca, repel = TRUE,col.ind = dataset$Salary_Level,
                col.var = "#2E9FDF", # Variables color
)

fviz_pca_biplot(res.pca, 
                col.ind = dataset$Salary_Level, palette = "jco", 
                addEllipses = TRUE, label = "var",
                col.var = "black", repel = TRUE,
                legend.title = "Laid Off") 

fviz_pca_biplot(res.pca, 
                # Fill individuals by groups
                geom.ind = "point",
                pointshape = 21,
                pointsize = 2.5,
                fill.ind = dataset$Salary_Level,
                col.ind = "black",
                # Color variable by groups
                legend.title = list(fill = "Laid Off", color = "Clusters"),
                repel = TRUE        # Avoid label overplotting
)+
  ggpubr::fill_palette("jco")+      # Indiviual fill color
  ggpubr::color_palette("npg")      # Variable colors

fviz_pca_biplot(res.pca, 
                # Individuals
                geom.ind = "point",
                fill.ind = dataset$Salary_Level, col.ind = "black",
                pointshape = 21, pointsize = 2,
                palette = "jco",
                addEllipses = TRUE,
                # Variables
                alpha.var ="contrib", col.var = "contrib",
                gradient.cols = "RdYlBu",
                
                legend.title = list(fill = "Laid Off", color = "Contrib",
                                    alpha = "Contrib")
)
```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```

```{r }

```


```{r }

```




```{r }

```



```{r }

```



```{r }

```



```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```